Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio

### 2018-07-06 paper/lostanlenRelevanceScattering.tex
brush up intro

An important problem is that manually analysing the recorded data to identify the quantities of interest is very costly. Some sort of pre-screening is therefore required to reduce the need for human expert listening and annotation.
To this aim, the most straightforward approach is to specify a closed set of sound classes, such as sounds classes expected to appear near the acoustic sensors. Computational models are then trained  for these classes which are used to automatically annotate recordings~\cite{quteprints96243}. A given time interval (\eg{} a single day) is then represented by the number of events detected during that interval for each class. This allows the scientist to drastically reduce the amount of information requiring manual processing. However, this approach has two drawbacks. First, it relies on trained models whose behavior on unseen data (such as different sensors) is prone to errors. Second, and more importantly, it is based on prior knowledge and thus cannot be considered for exploratory analysis where quantities of interest have yet to be defined.

To identify which parts need human inspection, we need tools that can detect both recurring patterns and sparsely distributed events. Identifying recurring patterns allows the user to focus on certain time points for manual annotation, while detection of more rare structures enables discovery of unforeseen phenomena.

With this aim, we need to design an algorithm for acoustic similarity retrieval, where the audio fragments judged ``most similar'' to a given query recording must be extracted from some larger dataset. To construct such an algorithm, we are required to represent an audio recording in a way that captures its distinctive qualities. A popular choice of representation is the bag of frames~\cite{aucouturier2007bag}, which describes an auditory scene recording using summary statistics of short-time features. Unfortunately, the bag-of-frames approach only captures the average structure of the scene, so the approach often fails when presented with highly dynamic scenes or those characterised by a few distinct sound events. Furthermore, experiments in cognitive psychology~\cite{dubois2006cognitive} and cognitive neuroscience~\cite{nelken2004processing} suggest that human acoustic perception is highly sensitive to such isolated sound events. We believe that the failure to model such distinct events is one of the reasons why the bag-of-frames representation is insufficient~\cite{lagrange:hal-01082501}.

@@

Caracterization of the similarity between audio recordings be they at the scale of the minute, the hour, the day or larger is of interest for many applications areas involving acoustic monitoring such as urban sound environment analysis and ecoacoustics. In this context, a popular approach is the bag of frames (BOF), first applied to the problem by Aucouturier et al.~\cite{aucouturier2007bag}.
It models an auditory scene using high-level summary statistics computed from local features, typically implemented by Gaussian mixture models (GMMs) of MFCCs.

>>>>

An important problem is that manually analysing the recorded data to identify the quantities of interest is very costly. Some sort of pre-screening is therefore required to reduce the need for human expert listening and annotation.
To this aim, the most straightforward approach is to specify a closed set of sound classes, such as sounds classes expected to appear near the acoustic sensors. Computational models are then trained  for these classes which are used to automatically annotate recordings~\cite{quteprints96243}. A given time interval (\eg{} a single day) is then represented by the number of events detected during that interval for each class. This allows the scientist to drastically reduce the amount of information requiring manual processing. However, this approach has two drawbacks. First, it relies on trained models whose prediction on unseen data -- \ie{} sensors outside of the training set -- is prone to errors. Secondly, and more importantly, it is based on prior knowledge and thus cannot be considered for exploratory analysis, in which quantities of interest have yet to be defined.

To identify which parts need human inspection, one needs tools that are able to detect both recurring patterns and sparsely distributed events. Identifying recurring patterns allows the user to focus on certain time points for manual annotation, while detection of more rare structures enables discovery of unforeseen phenomena.

With this aim, we need to design an algorithm for acoustic similarity retrieval, where the audio fragments judged ``most similar'' to a given query recording must be extracted from some larger dataset. To construct such an algorithm, we are required to represent an audio recording in a way that captures its distinctive qualities. A widespread choice of representation is the bag of frames~\cite{aucouturier2007bag}, which describes an auditory scene recording using summary statistics of short-time features. Unfortunately, the bag-of-frames approach only captures the average structure of the scene, so the approach often fails when presented with highly dynamic scenes or those characterised by a few distinct sound events. Furthermore, experiments in cognitive psychology~\cite{dubois2006cognitive} and cognitive neuroscience~\cite{nelken2004processing} suggest that human acoustic perception is highly sensitive to such isolated sound events. We believe that the failure to model such distinct events is one of the reasons why the bag-of-frames representation is insufficient~\cite{lagrange:hal-01082501}.

@@

Caracterization of the similarity between audio recordings be they at the scale of the minute, the hour, the day or larger is of interest for many applications areas involving acoustic monitoring such as urban sound environment analysis and ecoacoustics. In this context, a classical approach is the bag of frames (BOF), first applied to the problem by Aucouturier et al.~\cite{aucouturier2007bag}.
It models an auditory scene using high-level summary statistics computed from local features, typically implemented by Gaussian mixture models (GMMs) of MFCCs.
